Random Forests

### 2020-07-02 pkg/partykit/inst/ULGcourse-2020/slides_forest.Rnw
minor adaptions
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3034 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 666b75d7ae5b2ce734292dad650676a22614af71

{
  par(mfrow = c(2,2))
  set.seed(74)
  cf <- cforest(f, data = CPS1985, ntree = 2)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = predwage[,"predicted log(wage)"], type = 'l',
       ylab = "predicted log(wage)", xlab = "age", main = "2 trees")

  cf <- cforest(f, data = CPS1985, ntree = 4)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = predwage[,"predicted log(wage)"], type = 'l',
       ylab = "predicted log(wage)", xlab = "age", main = "4 trees")

  cf <- cforest(f, data = CPS1985, ntree = 20)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = predwage[,"predicted log(wage)"], type = 'l',
       ylab = "predicted log(wage)", xlab = "age", main = "20 trees")

  cf <- cforest(f, data = CPS1985, ntree = 200)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = predwage[,"predicted log(wage)"], type = 'l',
       ylab = "predicted log(wage)", xlab = "age", main = "200 trees")
}
@
\end{frame}

\begin{frame}
\frametitle{Variable importance}
{\bf Idea:} Asses the influence of a covariate on the model by evaluating the change in performance of the model (mean decrease in accuracy) after permutating the selected covariate.

>>>>

{
  par(mfrow = c(2,2))
  set.seed(74)
  cf <- cforest(f, data = CPS1985, ntree = 2)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = exp(predwage[,"predicted log(wage)"]), type = 'l',
       ylab = "predicted wage", xlab = "age", main = "Forest with 2 trees")

  cf <- cforest(f, data = CPS1985, ntree = 4)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = exp(predwage[,"predicted log(wage)"]), type = 'l',
       ylab = "predicted wage", xlab = "age", main = "Forest with 4 trees")

  cf <- cforest(f, data = CPS1985, ntree = 20)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = exp(predwage[,"predicted log(wage)"]), type = 'l',
       ylab = "predicted wage", xlab = "age", main = "Forest with 20 trees")

  cf <- cforest(f, data = CPS1985, ntree = 200)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = exp(predwage[,"predicted log(wage)"]), type = 'l',
       ylab = "predicted wage", xlab = "age", main = "Forest with 200 trees")
}
@
\end{frame}

\begin{frame}
\frametitle{Variable importance}
{\bf Problem:} In a single tree model the effect of each covariate can easily be investigated based on a simple illustration of the tree structure. However, in a forest model this would require to investigate each tree on its own including its specific setting (induced by random sampling).

{\bf Idea:} Asses the influence of a covariate on the forest model by evaluating the change in performance of the model (mean decrease in accuracy) after permutating the selected covariate.
